---
title: "Lab 2"
author: "Jeffrey Stevenson"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
EPAGAS <- read.csv("EPAGAS.csv")
head(EPAGAS)
```

## Task 3

```{r}
mpg <- EPAGAS$MPG
z <- (mpg-mean(mpg))/sd(mpg)
print(paste0("mean = ", round(mean(z), 4)), quote = 'FALSE')
print(paste0("variance = ", var(z)), quote = 'FALSE')
print(paste0("standard deviation = ", sd(z)), quote = 'FALSE')
```

Possible Outliers

```{r}
mpg[abs(z)>=2 & abs(z)<=3]
```

Definite Outliers

```{r}
mpg[abs(z)>3]
```

Plot

```{r}
library(lattice)
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 & abs(z)<=3,"Blue", "Black"))  
dotplot(mpg, col = mycol)
```

## Task 4

Boxplot

```{r}
boxplot(mpg, notch = TRUE, horizontal = TRUE, col = 'black', main='MPG Boxplot')
```

Proportion of data within 2 standard deviations of the mean of the mpg data

```{r}
k = 2
prop = 1-(1/k^2)
prop
```

Exact proportion within 2 standard deviation of the mean

```{r}
prop1 = length(mpg[abs(z)<=2])/length(mpg)
prop1
```

According to Chebysev's theorem, at least 75% of the data lies within 2 standard deviations of the mean. This agrees with the data, with 96% of the data lying withing 2 standard deviations of the mean.










