---
title: "Lab1"
author: "Jeffrey Stevenson"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

This is task 1 for Lab 1

```{r}
getwd()
```
## Task 2

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3

```{r}
ddt <- read.csv("DDT.csv")
names(ddt)
```

1. Qualitative Variables: River, Species

2. Quantitative Variables: Mile, Length, Weight, DDT

```{r}
ddt <- read.csv("DDT.csv")
summary(ddt)
```

3. Number of species in DDT dataset: 3 

4. 
```{r}
with(ddt, ddt[SPECIES == "LMBASS" & WEIGHT > 800,])
```

5.
```{r}
with(ddt, ddt[RIVER == "SCM" & DDT > 4.0,])
```

## Clicker Questions

Q1. Mean length of fish: 42.8125

```{r}
c(mean(ddt$LENGTH))
```

Q2. Standard Deviation of weight of fish: 376.5461

```{r}
c(sd(ddt$WEIGHT))
```

Q3. Is this the plot of LENGTH vs WEIGHT from the ddt data frame? 0 (No)

```{r}
with(ddt,plot(WEIGHT,LENGTH))
```

Q4. If v=1:20 what is the last value of v/20? 1

```{r}
v=1:20
v/20
```

## Task 4

1. Table of Rivers

```{r}
tab = table(ddt$RIVER)
print(tab)
```

2. Color Coded Barplot of Rivers

```{r}
barplot(tab, col=1:4)
```

3. Table of Rivers Crossed with Fish Species

```{r}
spriv=with(ddt,table(RIVER,SPECIES))
spriv
```

4. Barplot of Rivers Crossed with Fish Species
```{r}
barplot(spriv,beside=TRUE,col=1:4)
```

## Task 5

1. Pie Chart of Fish Species

```{r}
sp=with(ddt,table(SPECIES))
pie(sp,col=1:3)
```

2. Pie Chart of Rivers

```{r}
rv=with(ddt,table(RIVER))
pie(rv,col=1:4)
```

## Task 6

1. Boxplot of Variable DDT

```{r}
layout(matrix(c(1,2,3),nr=1,nc=3))# 1 row 3 cols
#layout.show(3)
with(ddt,boxplot(DDT,ylab="DDT",col="Blue",notch=TRUE))
```

2. Boxplot of Variable WEIGHT

```{r}
layout(matrix(c(1,2,3),nr=1,nc=3))# 1 row 3 cols
#layout.show(3)
with(ddt,boxplot(WEIGHT,ylab="WEIGHT",col="Red",notch=TRUE))
```

3. Boxplot of Variable LENGTH

```{r}
layout(matrix(c(1,2,3),nr=1,nc=3))# 1 row 3 cols
#layout.show(3)
with(ddt,boxplot(LENGTH,ylab="LENGTH",col="Green",notch=TRUE))
```

## Task 7

1. Coplot of LENGTH vs WEIGHT given RIVER

```{r}
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))
coplot(LENGTH~WEIGHT|RIVER,data=ddt, col=rivcol)
```

2. Coplot of DDT vs WEIGHT given SPECIES

```{r}
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(DDT~WEIGHT|SPECIES,data=ddt, col=fishcol)
```

## Task 8

```{r}
library(ggplot2)
g = ggplot(ddt, aes(x=SPECIES, y= WEIGHT, fill = RIVER) )
g = g + geom_boxplot() + ggtitle("Jeffrey Stevenson") + coord_equal(3/c(max(ddt$WEIGHT)))
g
```

```{r}
g = ggplot(ddt, aes(x=RIVER, y= LENGTH, fill = SPECIES) )
g = g + geom_violin() + ggtitle("Jeffrey Stevenson") + coord_equal(6/c(max(ddt$LENGTH)))
g
```

```{r}
g = ggplot(ddt, aes(x=WEIGHT, y= LENGTH, color = SPECIES) )
g = g + geom_point() + ggtitle("Jeffrey Stevenson") + coord_equal((c(max(ddt$WEIGHT))+1000)/c(max(ddt$LENGTH)))
g
```








