---
title: "Lab 4"
author: "Jeffrey Stevenson"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
spruce.df = read.csv("SPRUCE.csv")
tail(spruce.df)
```

## Task 3

```{r}
library(s20x)
trendscatter(Height~BHDiameter,f=0.5,data=spruce.df)

spruce.lm=with(spruce.df,lm(Height~BHDiameter))

height.res = residuals(spruce.lm)

height.fit = fitted(spruce.lm)

plot(height.fit,height.res)

trendscatter(height.fit,height.res)
```

**What shape is seen in the plot?**

The Residual vs. Fitted values plot shows a downward-facing parabola. Compared to the decreasingly positive-sloping Height vs. BHDiameter plot, it is clear
that this data does not fit a linear model well. 

**Residual plot using plot() function**

```{r}
plot(spruce.lm, which=1)
```

**Check normality**

```{r}
normcheck(spruce.lm,shapiro.wilk = TRUE)
```

**What is pvalue for Shapiro-Wilk test? What is NULL hypothesis in this case?**

The p-value is 0.29. Using an alpha value of 0.05, the NULL hypothesis (probability distribution of response y is normal) can be accepted. 

**Are residuals approximately normal in distribution with mean 0 and constant variance?**

```{r}
round(mean(height.res),6)
```

**Conclusions**

Considering that the data is normally distrubuted, and looking at the residuals vs. fitted values plot above, a straight line should not be applied to this data set. 

## Task 4

**Fit quadratic model to points**

```{r}
quad.lm = lm(Height~BHDiameter + I(BHDiameter^2), data=spruce.df)
```

**Height vs. BHDiameter with quadratic curve**

```{r}
plot(Height~BHDiameter,bg="Blue",pch=21,cex=1.2,
ylim=c(0,max(Height)),xlim=c(0,max(BHDiameter)), 
main="Spruce height prediction",data=spruce.df)

myplot=function(x){
 quad.lm$coeff[1] + quad.lm$coeff[2]*x + quad.lm$coeff[3]*x^2
}
curve(myplot, lwd=2, col="steelblue", add=TRUE)
```

**Make quad.fit**

```{r}
quad.fit = fitted(quad.lm)
```

**Make plot of residuals vs. fitted values**

```{r}
plot(quad.lm, which = 1)
```

**Construct QQ plot**

```{r}
normcheck(quad.lm, shapiro.wilk = TRUE)
```

**What is p-value? What do you conclude?**

The p-value is 0.684. Compared to an alpha of 0.05, the p-value is much greater. This means that we can accept the null hypothesis
that the data is normally distributed. The trendline in the residuals vs. fitted values plots indicated that this model fits well 
due to the trend line staying around 0 for the residual values.

## Task 5

**Summarize quad.lm**

```{r}
summary(quad.lm)
```

**Value of Bo, B1, and B2?**

$\hat{B_0} = 0.860896$
$\hat{B_1} = 1.469592$
$\hat{B_2} = -0.027457$

**Interval Estimates**

```{r}
ciReg(quad.lm)
```

**Equation of fitted line**

$\hat{B_0}+\hat{B_1}x+\hat{B_2}x^2=0.860896+1.469592x-0.027457x^2$

**Predict spruce height when diamter is 15, 18, and 20cm**

```{r}
predict(quad.lm, data.frame(BHDiameter=c(15,18,20)))

predict(spruce.lm, data.frame(BHDiameter=c(15,18,20)))
```

**Compare with previous predictions**

Overall, the quadratic predictions are greater than the linear predictions.

**Value of multiple $r^2$? Compare with previous model.**

```{r}
summary(quad.lm)$r.squared
summary(spruce.lm)$r.squared
```

The quadratic $r^2$ is 0.7741266, which is higher than the linear $r^2$ of 0.6569146. 

**Adjusted $r^2$**

```{r}
summary(quad.lm)$adj.r.squared
summary(spruce.lm)$adj.r.squared
```

The adjusted $r^2$ for the quadratic model is 0.7604373. The adjusted $r^2$ for the linear model is 0.6468239.
This means that the quadratic equation is a better fit because the value is closer to 1. 

**What does "Multiple $r^2$" mean in this case?**

Multiple $r^2$ shows how well the model represents the data. 

**Which model explains the most variability in height?**

```{r}
anova(quad.lm)
anova(spruce.lm)
```

The residual sum of squares for the quadratic model is less than that of the linear model, so 
the quadratic model fits the data better.

**TSS, MSS, RSS, and MSS/TSS**

```{r}
height.qfit=fitted(quad.lm)

RSS=with(spruce.df, sum((Height-height.qfit)^2))
RSS

MSS = with(spruce.df, sum((height.qfit-mean(Height))^2))
MSS

TSS = with(spruce.df, sum((Height-mean(Height))^2))
TSS

MSS/TSS
```

## Task 6

**Cook's plot**

```{r}
cooks20x(quad.lm)
```

The cooks distance is a measure of the influence a data point has on the regression model. 
This is typically used to identify outliers that negatively affect the regression model. 

**What does cooks distance for the quadratic model and data tell you?**

The cooks distance for this data and quadratic model shows that points 18, 21, and 24 are likely outliers
because of the influence they have on the model. 

**quad2.lm and summary**

```{r}
quad2.lm = lm(Height~BHDiameter + I(BHDiameter^2), data=spruce.df[-24,])
summary(quad2.lm)
```

**Compare with quad.lm**

```{r}
summary(quad.lm)
```

The multiple and adjusted $r^2$ values are larger for quad2.lm. The magnitude of the residuals are less 
for quad2.lm.

**Conclude**

Removing the outlier pointed out in the cooks plot significantly helped the quadratic model fit the data better.

## Task 7

**Proof**

First, we show 2 lines with point $x_k$ in common.
 $$l_1: y = B_0+B_1 x $$
$$l_2: y = B_0+\delta+(B_1+B_2)x $$
Next, we plug in $x_k$ and set the equations equal to each other because they share the same point.
 
$$ y_k=B_0+B_1x_k =B_0+\delta+(B_1+B_2)x_k $$
Next, redistribute the RHS.
$$ B_0+B_1x_k =B_0+\delta+B_1x_k+B_2x_k $$
$B_0$ and $B_1x_k$ cancel, leaving us with:

$$ 0=\delta+B_2x_k $$
Rearranged: 
$$\delta=-B_2x_k$$

Going back to $l_2$ for any $x$:
$$l_2: y = B_0+\delta+(B_1+B_2)x $$
Substitute $\delta=-B_2x_k$ into the equation:
$$l_2: y = B_0-B_2x_k+(B_1+B_2)x $$
Which can be rearranged to:
$$l_2: y = B_0+B_1x+B_2(x-x_k) $$
Now we have an equation that describes $l_2$ as an adjustment of $l_1$. 
We can use an indicator function $I(x>x_k)$ to allow our function to know where it should and should not include the adjustment. 
$$y = B_0+B_1x+B_2(x-x_k)I(x>x_k) $$
Where I() is 1 if $x>x_k$ and 0 else. 

**Reproduce Plot**

```{r}
sp2.df=within(spruce.df, X<-(BHDiameter-18)*(BHDiameter>18)) 
lmp=lm(Height~BHDiameter + X,data=sp2.df)
tmp=summary(lmp)
myf = function(x,coef){
  coef[1]+coef[2]*(x) + coef[3]*(x-18)*(x-18>0)
}
plot(spruce.df,main="Piecewise regression")
curve(myf(x,coef=tmp$coefficients[,"Estimate"] ),add=TRUE, lwd=2,col="Blue")
abline(v=18)
text(18,16,paste("R squared=",round(tmp$r.squared,4) ))
```

## Task 8

Takes a quantitative vector and returns a plot. This plot is produced according to the function.
```{r}
library(MATH4753stev0009)
myplot(1:40)
```





























